Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-01-31 CAV 2025 TODO/proofsb.tex
recover the file
update

{\color{blue}
\section{Parameter settings}

@@

We did not experiment with cutting planes (GCP-CROWN \cite{cutting}), as it needs an additional package, namely IBM CPLEX solver, we do not have access to. From \cite{cutting}, the number of undecided inputs of GCP-CROWN is $\leq 2\%$ better than $\alpha,\beta$-Crown on the DNNs we experimented with, far from the $10-40\%$ improvement seen from Hybrid MILP. The conclusion are thus unchanged.
}

@@

The setting for partial MILP for fully-connected DNNs is about how many neurons need to be opened (once set, the selection is automatic). The runtime depending crucially upon the number of open ReLU neurons, we set it quite tightly, only allowing few neuron deviation to accommodate to a particularly accurate/inaccurate bound computation (measure by the weight of the remaining Utility function). As complexity increases with the layer considered, as the size of the MILP model grows, we lower this number with the depth, only committing to an intermediate number for the output neuron (the number of output neurons  is smaller than hidden layer, and this is the most important computation). We experimentally set this number so that each computing the bounds in each hidden layer takes around the same time. Remember that in layer 1, partial MILP is not necessary and propagating bounds using interval arithmetic is already exact. We open [48,48] to compute bounds for hidden layer 2, [21,24] for layer 3, [11,14] for layer 4, [6,9] for layer 5, [3,6] for layer 6, [2,5] for layer 7, [1,4] for hidden layer 8 (if any), and we open [14,17] for the output layer.
{\color{blue} The exact number of open nodes in the range [a,a+3] is decided automatically for each neuron being computed : ReLUs are ranked according to their value by Utility, and the a top ReLUs are open. Then, ReLUs ranked a+1,a+2, a+3 are opened if their Utility value is larger than a small threshold. We set the threshold at 0.01. It should be seen as a way to save runtime when Utility knows that the next node by ranking (a+i) will not impact accuracy much (thanks to the upper bound from Proposition \ref{prop2}).}

@@

{\color{blue}
$\alpha,\beta$-Crown uses massively parallel ($>$4096 threads) GPU, while Partial MILP uses 20 CPU-threads.}

@@

{\color{blue}
Last, for Gurobi, we use a custom MIP-Gap (from $0.001$ to $0.1$) and time-out parameters, depending on the seen improvement and the possibility to make a node stable. This is low level implementation details that will be available in the code once the paper is accepted.
}

{\color{blue}

@@

We also provide in Table \ref{tab:example1} the raw numbers used to produce Figure \ref{fig_table3}.
Further, we tested with the SR \cite{BaB} and FSB heuristics \cite{FSB}, that chooses nodes to branch on for BaB (Branch and Bound). When SR and FSB are used to choose open nodes for pMILP, the accuracy is low as shown on Fig. \ref{fig_table3_new}: SR and FSB are worse than \cite{DivideAndSlide} for $<35$ open ReLU nodes, although unlike the latter, they can rank ReLU nodes in several layer before (which helps them a bit), and far worse than Utility. Further, FSB is performing worse than SR when choosing nodes for pMILP, while to choose nodes to branch on for BaB, it is the opposite \cite{FSB}. This likely means that the heuristic to choose nodes to branch for BaB is not adapted to choose nodes to open for pMILP.

\begin{table}[h!]	
	\centering
	\begin{tabular}{|c||c|c|c|c|c||c|c|c|c|}
		\hline
			&\multicolumn{5}{c||}{$X \subseteq$ Layer 2, max $=55$}&\multicolumn{4}{c|}{$X \subseteq$ Layers 1\&2, max $=116$} \\\cline{2-10}
		\text{$|X|$}  & \text{Random} & Huang &SR&FSB& \bf Utility& \text{Random} & SR & FSB &\bf Utility\\ \hline
		\hline
		0  (LP) &   1.761  &1.761&1.761&1.761& 1.761 & 1.761  &  1.761& 1.761 & 1.761\\ \hline \hline
		5  &   1.729&1.704 &1.6933&1.6934& 1.603&  1.729& 1.6533 & 1.6574 & \bf 1.532 \\ \hline
		10  &  1.701 &  1.651 &1.6432&1.6393&1.517& 1.696 &  1.5672& 1.5741 & \bf  1.371\\ \hline
		15  & 1.671 &  1.599 &1.6061& 1.5996&1.466&  1.653& 1.4954& 1.5007 & \bf  1.247\\ \hline
		20  &  1.635 & 1.557&1.5730&1.5674&1.438 & 1.619 &1.4321 & 1.4385 & \bf  1.145\\ \hline
		25  &  1.601 & 1.519 &1.5431&1.5405&1.427&  1.586 &1.3783 & 1.3823 & \bf 1.061\\ \hline
		30  & 1.574 & 1.489 &1.5129&1.5151&1.425& 1.546 &  1.3328& 1.3330 & \bf  0.989 \\ \hline
		35  &  1.542 & 1.465&1.4871&1.4899&1.424 & 1.502 & 1.2891 & 1.2934 & \bf 0.934 \\ \hline
		40  & 1.512 & 1.447 &1.4659&1.4686&1.424& 1.469 & 1.2574& 1.2630 & \bf 0.921 \\ \hline \hline
		max & 1.424  &1.424&1.424&1.424& 1.424 & 0.895 & 0.895 & 0.895 & 0.895   \\ \hline
	\end{tabular}
	\caption{Average uncertainty of $\MILP_X$ for nodes of the third layer, with ReLU nodes of the (1st and) 2nd layer, chosen by our {\bf Utility} function vs \cite{DivideAndSlide} vs vs SR vs FSB vs random.}
	\label{tab:example1}
	%\vspace{-0.6cm}
\end{table}

	\begin{figure}[h!]
		\centering
		\vspace*{-0.3cm}
		\includegraphics[height=9cm]{New Plot for table 3}.
		\vspace*{-0.4cm}
		\caption{Average uncertainty of $\MILP_X$ for nodes of the third layer, with ReLU nodes of the (1st and) 2nd layer, chosen by our {\bf Utility} function vs \cite{DivideAndSlide} vs vs SR vs FSB vs random.}
		\label{fig_table3_new}
	\end{figure}

@@

\begin{table}[b!]
	\centering
	\begin{tabular}{||l||c|c||c||}
		\hline \hline
		Network & $\alpha,\beta$-Crown & Refined $\beta$-Crown & PRIMA \\ 		  
		\hline
		MNIST $5 \times 100$ & N/A  & 14.3\% (102s) & 33.2\% (159s)\\ \hline
		MNIST $5 \times 200$ & N/A & 13.7\% (86s) & 21.1\% (224s) \\ \hline
		MNIST $8 \times 100$ & N/A  & 20.0\% (103s) & 39.2\% (301s)   \\ \hline
		MNIST $8 \times 200$ & N/A & 17.6\% (95s) & 28.7\% (395s)  \\ \hline
		MNIST $6 \times 500$ & 51\% (16s) & $-$ & 64\% (117s) \\ \hline
		CIFAR CNN-B-adv & 18.5\% (32s) & $-$ & 27\% (344s)\\ \hline \hline
		CIFAR ResNet & 0\% (2s) & $-$ & 0\% (2s) \\ \hline \hline
	\end{tabular}
	\caption{Undecided images ($\%$, {\em lower is better}), as computed by $\alpha,\beta$-Crown, Refined $\beta$-Crown, and PRIMA, as reported in \cite{crown}, except for $6 \times 500$ that we run ourselves. N/A means that \cite{crown} did not report the numbers, while $-$ means that Refined $\beta$-Crown cannot be run on these DNNs.}
	\label{table9}
	%\begin{tablenotes}
	%	\footnotesize
%		\item Most data is directly from \cite{crown}. N/A means no data either in \cite{crown} or by our running.
%		\item  $^*$ The data in this row is from our own running on first 100 images of the MNIST dataset.
%		\item  $^{**}$ The data is from \cite{crown} on first 200 images of the CIFAR10 dataset.
%	\end{tablenotes}
	\end{table}

@@

{\color{blue}

>>>>

\section{Parameter settings}

@@

We did not experiment with cutting planes (GCP-CROWN \cite{cutting}), as it needs an additional package, namely IBM CPLEX solver, we do not have access to. From \cite{cutting}, the number of undecided inputs of GCP-CROWN is $\leq 2\%$ better than $\alpha,\beta$-Crown on the DNNs we experimented with, far from the $10-40\%$ improvement seen from Hybrid MILP. The conclusion are thus unchanged.

@@

The setting for partial MILP for fully-connected DNNs is about how many neurons need to be opened (once set, the selection is automatic). The runtime depending crucially upon the number of open ReLU neurons, we set it quite tightly, only allowing few neuron deviation to accommodate to a particularly accurate/inaccurate bound computation (measure by the weight of the remaining Utility function). As complexity increases with the layer considered, as the size of the MILP model grows, we lower this number with the depth, only committing to an intermediate number for the output neuron (the number of output neurons  is smaller than hidden layer, and this is the most important computation). We experimentally set this number so that each computing the bounds in each hidden layer takes around the same time. Remember that in layer 1, partial MILP is not necessary and propagating bounds using interval arithmetic is already exact. We open [48,48] to compute bounds for hidden layer 2, [21,24] for layer 3, [11,14] for layer 4, [6,9] for layer 5, [3,6] for layer 6, [2,5] for layer 7, [1,4] for hidden layer 8 (if any), and we open [14,17] for the output layer.
 The exact number of open nodes in the range [a,a+3] is decided automatically for each neuron being computed : ReLUs are ranked according to their value by Utility, and the a top ReLUs are open. Then, ReLUs ranked a+1,a+2, a+3 are opened if their Utility value is larger than a small threshold. We set the threshold at 0.01. It should be seen as a way to save runtime when Utility knows that the next node by ranking (a+i) will not impact accuracy much (thanks to the upper bound from Proposition \ref{prop2}).

@@

$\alpha,\beta$-Crown uses massively parallel ($>$4096 threads) GPU, while Partial MILP uses 20 CPU-threads.

@@

Last, for Gurobi, we use a custom MIP-Gap (from $0.001$ to $0.1$) and time-out parameters, depending on the seen improvement and the possibility to make a node stable. This is low level implementation details that will be available in the code once the paper is accepted.

@@

%
%We also provide in Table \ref{tab:example1} the raw numbers used to produce Figure \ref{fig_table3}.
%Further, we tested with the SR \cite{BaB} and FSB heuristics \cite{FSB}, that chooses nodes to branch on for BaB (Branch and Bound). When SR and FSB are used to choose open nodes for pMILP, the accuracy is low as shown on Fig. \ref{fig_table3_new}: SR and FSB are worse than \cite{DivideAndSlide} for $<35$ open ReLU nodes, although unlike the latter, they can rank ReLU nodes in several layer before (which helps them a bit), and far worse than Utility. Further, FSB is performing worse than SR when choosing nodes for pMILP, while to choose nodes to branch on for BaB, it is the opposite \cite{FSB}. This likely means that the heuristic to choose nodes to branch for BaB is not adapted to choose nodes to open for pMILP. 
%
%\begin{table}[h!]	
%	\centering
%	\begin{tabular}{|c||c|c|c|c|c||c|c|c|c|}
%		\hline
%			&\multicolumn{5}{c||}{$X \subseteq$ Layer 2, max $=55$}&\multicolumn{4}{c|}{$X \subseteq$ Layers 1\&2, max $=116$} \\\cline{2-10}
%		\text{$|X|$}  & \text{Random} & Huang &SR&FSB& \bf Utility& \text{Random} & SR & FSB &\bf Utility\\ \hline
%		\hline
%		0  (LP) &   1.761  &1.761&1.761&1.761& 1.761 & 1.761  &  1.761& 1.761 & 1.761\\ \hline \hline
%		5  &   1.729&1.704 &1.6933&1.6934& 1.603&  1.729& 1.6533 & 1.6574 & \bf 1.532 \\ \hline
%		10  &  1.701 &  1.651 &1.6432&1.6393&1.517& 1.696 &  1.5672& 1.5741 & \bf  1.371\\ \hline
%		15  & 1.671 &  1.599 &1.6061& 1.5996&1.466&  1.653& 1.4954& 1.5007 & \bf  1.247\\ \hline
%		20  &  1.635 & 1.557&1.5730&1.5674&1.438 & 1.619 &1.4321 & 1.4385 & \bf  1.145\\ \hline
%		25  &  1.601 & 1.519 &1.5431&1.5405&1.427&  1.586 &1.3783 & 1.3823 & \bf 1.061\\ \hline
%		30  & 1.574 & 1.489 &1.5129&1.5151&1.425& 1.546 &  1.3328& 1.3330 & \bf  0.989 \\ \hline
%		35  &  1.542 & 1.465&1.4871&1.4899&1.424 & 1.502 & 1.2891 & 1.2934 & \bf 0.934 \\ \hline
%		40  & 1.512 & 1.447 &1.4659&1.4686&1.424& 1.469 & 1.2574& 1.2630 & \bf 0.921 \\ \hline \hline
%		max & 1.424  &1.424&1.424&1.424& 1.424 & 0.895 & 0.895 & 0.895 & 0.895   \\ \hline
%	\end{tabular}
%	\caption{Average uncertainty of $\MILP_X$ for nodes of the third layer, with ReLU nodes of the (1st and) 2nd layer, chosen by our {\bf Utility} function vs \cite{DivideAndSlide} vs vs SR vs FSB vs random.}
%	\label{tab:example1}
%	%\vspace{-0.6cm}
%\end{table}
%
%
%
%	\begin{figure}[h!]
%		\centering
%		\vspace*{-0.3cm}
%		\includegraphics[height=9cm]{New Plot for table 3}.
%		\vspace*{-0.4cm}
%		\caption{Average uncertainty of $\MILP_X$ for nodes of the third layer, with ReLU nodes of the (1st and) 2nd layer, chosen by our {\bf Utility} function vs \cite{DivideAndSlide} vs vs SR vs FSB vs random.}
%		\label{fig_table3_new}
%	\end{figure}

@@

\begin{table}[b!]
	\centering
	\begin{tabular}{||l||c|c||c||}
		\hline \hline
		Network & $\alpha,\beta$-Crown & Refined $\beta$-Crown & PRIMA \\ 		  
		\hline
		MNIST $5 \times 100$ & N/A  & 14.3\% (102s) & 33.2\% (159s)\\ \hline
		MNIST $5 \times 200$ & N/A & 13.7\% (86s) & 21.1\% (224s) \\ \hline
		MNIST $8 \times 100$ & N/A  & 20.0\% (103s) & 39.2\% (301s)   \\ \hline
		MNIST $8 \times 200$ & N/A & 17.6\% (95s) & 28.7\% (395s)  \\ \hline
		MNIST $6 \times 500$ & 51\% (16s) & $-$ & 64\% (117s) \\ \hline
		CIFAR CNN-B-adv & 18.5\% (32s) & $-$ & 27\% (344s)\\ \hline \hline
		CIFAR ResNet & 0\% (2s) & $-$ & 0\% (2s) \\ \hline \hline
	\end{tabular}
	\caption{Undecided images ($\%$, {\em lower is better}), as computed by $\alpha,\beta$-Crown, Refined $\beta$-Crown, and PRIMA, as reported in \cite{crown}, except for $6 \times 500$ that we run ourselves. N/A means that \cite{crown} did not report the numbers, while $-$ means that Refined $\beta$-Crown cannot be run on these DNNs.}
	\label{table9}
	\begin{tablenotes}
		\footnotesize
		\item Most data is directly from \cite{crown}. N/A means no data either in \cite{crown} or by our running.
		\item  $^*$ The data in this row is from our own running on first 100 images of the MNIST dataset.
		\item  $^{**}$ The data is from \cite{crown} on first 200 images of the CIFAR10 dataset.
	\end{tablenotes}
	\end{table}

@@
